---
title: "Introduction to the Spatstat Package"
author: "Zhongyu Zhou"
date: "2024-10-09"
output:
  ioslides_presentation:
    widescreen: true
    self_contained: true
    css: |
      body {
        background-color: white;
      }
      .title-slide h1, .title-slide h2 {
        text-align: center;
        color: black;
      }
      .slide h1, .slide h2, .slide h3 {
        text-align: center;
        color: black;
      }
---
# Introduction to `spatstat`

`spatstat` is an R package used for spatial point pattern analysis. It provides a comprehensive set of tools for analyzing and visualizing spatial point patterns, performing cluster analysis, estimating intensity, and simulating point processes. The package is highly useful in a wide range of fields, such as ecology, epidemiology, and geography.

For more information, visit the [official CRAN page](https://cran.r-project.org/web/packages/spatstat/index.html).

# Author Information

One of the primary authors of `spatstat` is Adrian Baddeley. He is a prominent figure in the field of spatial statistics and has contributed to several other R packages related to spatial data analysis. Adrian Baddeley is affiliated with Curtin University, Australia, and has authored multiple research papers in the domain of spatial analysis and point processes.

# Features Overview

- **Visualization**: Plot point patterns and various diagnostic plots to understand the spatial structure of data.
- **Spatial Analysis**: Perform functions like K-function analysis and G-function to evaluate clustering or regularity.
- **Density Estimation**: Estimate the density of points over a spatial region.
- **Simulation**: Simulate random point patterns based on different spatial processes (e.g., Poisson process).
- **Covariate Analysis**: Fit point process models using covariate data.

In the following sections, a simple demonstration of some key functionalities will be provided.

# Demonstration of Code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, echo=TRUE}
# Load required packages
library(spatstat)
```

## Data Preparation

We will load sample data and convert it into a point pattern object.

```{r data-preparation, echo=TRUE}
# Generate some sample data
set.seed(42)
x <- runif(100, min = 0, max = 1)
y <- runif(100, min = 0, max = 1)

# Define the window and create a point pattern object
window <- owin(c(0, 1), c(0, 1))
locations <- ppp(x, y, window = window)

# Plot the data
plot(locations, main = "Sample Point Pattern")
```

## K-function Analysis

The K-function helps to evaluate whether the points exhibit clustering or regularity.

```{r k-function-analysis, echo=TRUE}
# Compute and plot the K-function
K <- Kest(locations)
plot(K, main = "K-function Analysis")
```

## Density Estimation

Estimate the density of points in a given region to understand the hotspots.

```{r density-estimation, echo=TRUE}
# Estimate density and plot the results
density_map <- density(locations)
plot(density_map, main = "Density Estimation")
```

## Point Pattern Simulation

Simulate a random point pattern using a homogeneous Poisson process.

```{r point-pattern-simulation, echo=TRUE}
# Simulate random point pattern
random_pattern <- rpoispp(100, win = window)

# Plot the simulated pattern
plot(random_pattern, main = "Simulated Point Pattern")
```

# Summary

In this presentation, we introduced the `spatstat` package, which is an essential tool for spatial point pattern analysis. We demonstrated some of the key features, including point pattern visualization, K-function analysis, density estimation, and point pattern simulation.

The `spatstat` package is highly useful for spatial data analysis across various domains and can help researchers understand the underlying patterns in spatial data more effectively.

Thank you for your attention!