---
title: "Spatstat Case"
author: "Zhongyu Zhou"
date: "2024-10-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spatstat)
library(OpenStreetMap)
```

# Data Preparation

First, we will load the weather station data and convert it into a point pattern data object in `spatstat`.

```{r data-preparation}
# Load data
your_data <- read.csv("data/annual_conc_by_monitor_2023.csv")

# Select longitude and latitude columns
your_data <- your_data[, c("Longitude", "Latitude")]

# Remove duplicate points
your_data <- your_data[!duplicated(your_data), ]

# Define the point pattern window range
window <- owin(xrange = range(your_data$Longitude, na.rm = TRUE),
yrange = range(your_data$Latitude, na.rm = TRUE))

# Convert the data to a point pattern object
locations <- ppp(x = your_data$Longitude, y = your_data$Latitude, window = window)

# Plot point pattern with OSM basemap
base_map <- openmap(upperLeft = c(max(your_data$Latitude), min(your_data$Longitude)), lowerRight = c(min(your_data$Latitude), max(your_data$Longitude)), type = "osm")
base_map_projected <- openproj(base_map, projection = "+proj=longlat +datum=WGS84")

plot(base_map_projected, main = "Meteorological Stations in US")
points(your_data$Longitude, your_data$Latitude, pch = 20, col = "blue", cex = 0.6)
```

# K-function Analysis

The K-function is used to evaluate the spatial clustering between points in a point pattern, helping to determine whether the points are randomly distributed, clustered, or evenly distributed.

```{r k-function}
# Calculate K function
K <- Kest(locations)

# Plot K function
plot(K, main = "K-function Analysis of Meteorological Stations")

# The black curve(spatial clustering patterns of actual data) is significantly higher than the blue completely random distribution (CSR) curve, which indicates that the distribution of weather stations has significant spatial aggregation. That is, weather stations are not randomly distributed but tend to be concentrated in certain areas.

# The red and green curves are the edge-corrected K functions, which reduce the deviation due to boundary effects.
```

# Density Estimation

Density estimation can be used to observe the distribution density of meteorological stations in space and help identify hot spots.

```{r density-estimation}
# Estimate the density of points
density_map <- density(locations)

# Plot the density map
plot(density_map, main = "Density Estimation of Meteorological Stations")
```

# Point Pattern Simulation

Simulate a new point pattern using a completely random distribution (CSR) and compare with the actual data.

```{r point-simulation}
# Simulate a completely random point pattern, generate about 300 points
random_pattern <- rpoispp(300 / area.owin(window), win = window)

# Plot the simulated point pattern
plot(base_map_projected, main = "Simulated Random Point Pattern with Basemap")
points(random_pattern$x, random_pattern$y, pch = 20, col = "red", cex = 0.6)
```

# Multi-type Point Pattern Analysis

If the weather stations have different categories (e.g. different types of weather stations), we can perform multi-type point pattern analysis.

```{r multi-type}
# Make sure the data still has enough categories after removing duplicate points
your_data$station_type <- sample(c("Type A", "Type B"), nrow(your_data), replace = TRUE)

# Create a multi-type point pattern object
multi_type_data <- ppp(x = your_data$Longitude, y = your_data$Latitude, marks = factor(your_data$station_type), window = window)

# Plot the multi-type point pattern
plot(multi_type_data, main = "Multi-type Meteorological Stations")

# Calculate the multi-type K-function
multi_K <- Kcross(multi_type_data, i = "Type A", j = "Type B")
plot(multi_K, main = "Multi-type K-function Analysis")

```